Predicting the Outcome of Patient-Provider Communication Sequences using Recurrent Neural Networks and Probabilistic Models

### 2018-01-11 paper-amia-summit-2018/amia_summit_2018.tex
update reviewer's comments and manuscript for camera-ready version

\textbf {Recurrent Neural Networks (RNN)} are a class of neural networks that have an internal memory, which makes them particularly suitable for processing sequences of observations. The ability of RNNs to capture long-term dependencies and remember past observations for predicting future observations is their main advantage over MCs and HMMs. These features are very useful in analysis of motivational interviews, in which any behavior observed at a particular point in the interview may be indicative of other behaviors that are observed later. In order to mitigate the vanishing gradient problem of earlier versions of RNN \cite{bengio1993problem}, Hochreiter et al.\cite{hochreiter1997long} proposed Long Short Term Memory networks (LSTM). There are several variants of LSTM model, among which the most notable one is the Gated Recurrent Unit\cite{cho2014properties} (GRU). GRUs are simpler than LSTMs and have been shown to be effective for a variety of Natural Language Processing tasks \cite{cho2014properties}. GRU is formally defined as follows:
\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\label{eq:resetgru}
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\label{eq:candidategru}
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product. The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors. The activation $h_t$ of the GRU at time $t$ is a linear combination of the previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).

>>>>

\textbf {Recurrent Neural Networks (RNN)} are a class of neural networks that have an internal memory, which makes them particularly suitable for processing sequences of observations. The ability of RNNs to capture long-term dependencies and remember past observations for predicting future observations is their main advantage over MCs and HMMs. These features are very useful in the analysis of motivational interviews, in which any behavior observed at a particular point in the interview may be indicative of other behaviors that are observed later. In order to mitigate the vanishing gradient problem of earlier versions of RNN \cite{bengio1993problem}, Hochreiter et al.\cite{hochreiter1997long} proposed Long Short Term Memory networks (LSTM). There are several variants of LSTM model, among which the most notable one is the Gated Recurrent Unit\cite{cho2014properties} (GRU). GRUs are simpler than LSTMs and have been shown to be effective for a variety of Natural Language Processing tasks \cite{cho2014properties}. GRU is formally defined as follows:
\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\label{eq:resetgru}
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\label{eq:candidategru}
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product. The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors. The activation $h_t$ of the GRU at time $t$ is a linear combination of the previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
